---
title: "Code for analysis"
output: html_notebook
---

***

The following script reproduces the main analysis of the paper "Do the two make a pair? The environmental effects of the green and digital “twin” transition in European regions". For further questions and/or clarifications: s.bianchini@unistra.fr 


***
### Data 

```{r}
 rm(list = ls())

 library(dplyr)
 library(Hmisc)
 library(AER)
 library(corrplot)
 library(lmtest)
 library(sandwich)
 library(stargazer)
 library(PerformanceAnalytics)

 load("~/Desktop/data.RData")
```


***
### Descriptive statistics
See Table B1 - APPENDIX B. Additional statistics

```{r}
# Select variables
 df <- dat %>%
   select(ghg_log, psm_ghg_log, gt, dt, 
          add_manuf, ai, big_data, comp_infra, iot, robotics,
          pop, va_ind, urban, intermediate, rural, 
          rile, extreme_left, qoi) 
 
# Table with statistics
 stargazer(as.data.frame(df), 
           covariate.labels = c("GHG Emissions (log)", "Pre-sample Mean GHG (log)", 
                                "Green Tech", "Digital Tech", "Additive Manufacturing", "AI", "Big Data",
                                "Computing Infrastructures",  "IoT", "Robotics", "Population", 
                                "Value Added Manufacturing", "Urban", "Intermediate", 
                                "Rural" , "RILE", "Extreme Left","Quality of Institutions"), 
           align=TRUE,
           type="text")
```

***
### Correlations
See Figure B2 - APPENDIX B. Additional statistics
```{r}
# Select variables
 df <- dat %>%
   select(ghg_log, psm_ghg_log, gt_log_avg, dt_log_avg, 
          pop_log, va_ind_log, urban, intermediate, rural, 
          rile, extreme_left, qoi_log) 

# Compute correlation coefficients
 correlation.matrix <- cor(df)
 
# Results - Table
 # stargazer(correlation.matrix, align=TRUE, type="text")
 
 colnames(correlation.matrix) <- c("GHG Emissions (log)", "Pre-sample Mean GHG (log)", "Green Tech. (log)",
                                   "Digital Tech (log)", "Population (log)", 
                                   "Value Added Manufacturing (log)", "Urban", "Intermediate", "Rural",
                                   "RILE", "Extreme Left", "Quality of Institutions (log)")
 
 rownames(correlation.matrix) <- c("GHG Emissions (log)", "Pre-sample Mean GHG (log)", "Green Tech. (log)",
                                   "Digital Tech (log)", "Population (log)", 
                                   "Value Added Manufacturing (log)", "Urban", "Intermediate", "Rural", 
                                   "RILE", "Extreme Left", "Quality of Institutions (log)")
# Results - Heatmap
 corrplot(correlation.matrix, method = 'square',  diag = F, tl.col = "black", tl.cex = 0.8, addCoef.col ='black', number.cex = 0.7)
```

***
### Tobit estimates
See Table 2. Tobit estimates – The effect of green and digital endowment on regional emissions 
```{r}
# Estimate the models
 mod1 <- tobit(ghg_log ~ gt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 mod2 <- tobit(ghg_log ~ gt_log_avg + dt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 mod3 <- tobit(ghg_log ~ gt_log_avg + dt_log_avg + I(gt_log_avg * dt_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)

# Compute robust standard errors
 mod1 <- coeftest(mod1, vcov.=vcovCL(mod1, cluster=rownames("metro_code"), type="HC0"))
 mod2 <- coeftest(mod2, vcov.=vcovCL(mod2, cluster=rownames("metro_code"), type="HC0"))
 mod3 <- coeftest(mod3, vcov.=vcovCL(mod3, cluster=rownames("metro_code"), type="HC0"))
 
# Results
# Note: Re-run the above code without calculating robust standard errors to obtain information on nb. observations, log-likelihood, Wald tests, etc. 
 stargazer(mod1, mod2, mod3,
           covariate.labels = c("Green Tech (log)", "Digital Tech (log)", "Green x Digital Tech (log)", 
                                "Population (log)", "Value Added Manuf (log)", "Intermediate", "Rural",
                                "Pre-sample Mean GHG (log)"),
           omit = c("year", "Constant"),
           dep.var.labels = "GHG Emissions (log)",
           align=TRUE,
           object.names = TRUE, 
           df = FALSE,
           type="text")
```


***
### Tobit estimates - digital categories
See Table 4. Tobit estimates – The effect of digital technology categories on regional emissions

```{r}
# Estimate the models
 mod1  <- tobit(ghg_log ~ gt_log_avg + add_manuf_log_avg + I(gt_log_avg*add_manuf_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, data = dat)
 mod2  <- tobit(ghg_log ~ gt_log_avg + ai_log_avg + I(gt_log_avg*ai_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, data = dat)
 mod3  <- tobit(ghg_log ~ gt_log_avg + big_data_log_avg + I(gt_log_avg*big_data_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, data = dat)
 mod4  <- tobit(ghg_log ~ gt_log_avg + comp_infra_log_avg + I(gt_log_avg*comp_infra_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, data = dat)
 mod5  <- tobit(ghg_log ~ gt_log_avg + iot_log_avg + I(gt_log_avg*iot_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, data = dat)
 mod6  <- tobit(ghg_log ~ gt_log_avg + robotics_log_avg + I(gt_log_avg*robotics_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, data = dat)
 
# Compute robust standard errors
 mod1 <- coeftest(mod1, vcov.=vcovCL(mod1, cluster=rownames("metro_code"), type="HC0"))
 mod2 <- coeftest(mod2, vcov.=vcovCL(mod2, cluster=rownames("metro_code"), type="HC0"))
 mod3 <- coeftest(mod3, vcov.=vcovCL(mod3, cluster=rownames("metro_code"), type="HC0"))
 mod4 <- coeftest(mod4, vcov.=vcovCL(mod4, cluster=rownames("metro_code"), type="HC0"))
 mod5 <- coeftest(mod5, vcov.=vcovCL(mod5, cluster=rownames("metro_code"), type="HC0"))
 mod6 <- coeftest(mod6, vcov.=vcovCL(mod6, cluster=rownames("metro_code"), type="HC0"))
 
# Results
# Note: Re-run the above code without calculating robust standard errors to obtain information on nb. observations, log-likelihood, Wald tests, etc. 
 stargazer(mod1, mod2, mod3, mod4, mod5, mod6,
           covariate.labels = c("Green Tech (log)", 
                                
                                "Additive Manuf (log)",
                                "Green x Additive Manuf (log)",
                                
                                "AI (log)", 
                                "Green x AI (log)",
                                
                                "Big Data (log)", 
                                "Green x Big Data (log)",
                                
                                "Computing Infra (log)", 
                                "Green x Computing Infra (log)",
                                
                                "IoT (log)", 
                                "Green x IoT (log)",
                                
                                "Robotics (log)", 
                                "Green x Robotics (log)", 
                                
                                "Population (log)", "Value Added Manuf (log)",
                                "Intermediate", "Rural", 
                                "Pre-sample Mean GHG (log)"),
           omit = c("year", "Constant"),
           dep.var.labels = "GHG Emissions (log)",
           align=TRUE,
           object.names = TRUE, 
           df = FALSE,
           type="text")
```


***
### Instrumental variables (first step)
See Table B4. First step IV estimates 

```{r} 
#-- Estimation of reduced forms (3 configurations with different sets of instruments) --
 
# Instruments: qoi_log + rile
 reduced.form1 <- lm(gt_log_avg ~ qoi_log + rile + dt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 
# Instruments: qoi_log + extreme_left
 reduced.form2 <- lm(gt_log_avg ~ qoi_log + extreme_left + dt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 
# Instruments: qoi_log + extreme_left + extreme_right
 reduced.form3 <- lm(gt_log_avg ~ qoi_log + extreme_left + extreme_right + dt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 
# Compute fitted values (i.e, instrumented variable)
 dat$gt_log_avg_pred1 <- reduced.form1$fitted.values
 dat$gt_log_avg_pred2 <- reduced.form2$fitted.values
 dat$gt_log_avg_pred3 <- reduced.form3$fitted.values
 
# Compute F-test on instruments
 linearHypothesis(reduced.form1, c("qoi_log=0", "rile=0"))
```

```{r} 
 linearHypothesis(reduced.form2, c("qoi_log=0", "extreme_left=0"))
```

```{r} 
 linearHypothesis(reduced.form3, c("qoi_log=0", "extreme_left=0", "extreme_right=0"))
```

```{r}  
# Compute robust standard errors
 reduced.form1 <- coeftest(reduced.form1, vcov.=vcovCL(reduced.form1, cluster=rownames("metro_code"), type="HC0"))
 reduced.form2 <- coeftest(reduced.form2, vcov.=vcovCL(reduced.form2, cluster=rownames("metro_code"), type="HC0"))
 reduced.form3 <- coeftest(reduced.form3, vcov.=vcovCL(reduced.form3, cluster=rownames("metro_code"), type="HC0"))
 
# Results
# Note: Re-run the above code without calculating robust standard errors to obtain information on nb. observations, log-likelihood, Wald tests, etc. 
 stargazer(reduced.form1, reduced.form2, reduced.form3,
           covariate.labels = c("QOI (log)", "RILE", "Extreme Left", "Extreme Right", "Digital Tech (log)", 
                                "Population (log)", "Value Added Manuf (log)", "Intermediate", "Rural",
                                "Pre-sample Mean GHG (log)"),
           omit = c("year", "Constant"),
           dep.var.labels = "Green Tech. (log)",
           align=TRUE,
           object.names = TRUE, 
           df = FALSE,
           type="text") 
```


***
### Instrumental variables (second step)
Table 3. IV Estimates – The effect of green and digital endowment on regional emissions

```{r}  
# Note: 'gt_log_avg_pred2' is the instrumented variable that passes all the tests 
 
# Estimate the models
 mod1_2step <- lm(ghg_log ~ gt_log_avg_pred2 + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 mod2_2step <- lm(ghg_log ~ gt_log_avg_pred2 + dt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 mod3_2step <- lm(ghg_log ~ gt_log_avg_pred2 + dt_log_avg + I(gt_log_avg_pred2 * dt_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 
# Compute robust standard errors
 mod1_2step <- coeftest(mod1_2step, vcov.=vcovCL(mod1_2step, cluster=rownames("metro_code"), type="HC0"))
 mod2_2step <- coeftest(mod2_2step, vcov.=vcovCL(mod2_2step, cluster=rownames("metro_code"), type="HC0"))
 mod3_2step <- coeftest(mod3_2step, vcov.=vcovCL(mod3_2step, cluster=rownames("metro_code"), type="HC0"))
 
# Results
# Note: Re-run the above code without calculating robust standard errors to obtain information on nb. observations, log-likelihood, Wald tests, etc. 
 stargazer(mod1_2step, mod2_2step, mod3_2step, 
           covariate.labels = c("Green Tech (log) [Instrumented]", "Digital Tech (log)", 
                                "Green [Instr.] x Digital Tech (log)",
                                "Population (log)", "Value Added Manuf (log)", "Intermediate", "Rural", 
                                "Pre-sample Mean GHG (log)"),
           omit = c("year", "Constant"),
           align=TRUE,
           object.names = TRUE, 
           df = FALSE,
           type="text") 
```

```{r}  
# Over-identification tests  
# Note: Manually compute the chi-square statistics and the associated p-values
 
#--- Mod1 

# Reduced form and residuals
 reduced.form2 <- lm(gt_log_avg ~ qoi_log + extreme_left + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)

 dat$gt_log_avg_pred2 <- reduced.form2$fitted.values

# Chi-square and p-value  
 mod1_2step <- lm(ghg_log ~ gt_log_avg_pred2 + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 dat$residuals <- mod1_2step$residuals
 chisq <- summary(lm(residuals ~ qoi_log + extreme_left, dat))$r.squared * nrow(dat)
 chisq
 pchisq(chisq, 1, lower.tail=FALSE)
```

```{r}  
#--- Mod2

# Reduced form and residuals (same for Mod2 and Mod3)
 reduced.form2 <- lm(gt_log_avg ~ qoi_log + extreme_left + dt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)

 dat$gt_log_avg_pred2 <- reduced.form2$fitted.values
  
# Chi-square and p-value  
 mod2_2step <- lm(ghg_log ~ gt_log_avg_pred2 + dt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 dat$residuals <- mod2_2step$residuals
 chisq <- summary(lm(residuals ~ qoi_log + extreme_left, dat))$r.squared * nrow(dat)
 chisq
 pchisq(chisq, 1, lower.tail=FALSE)
```

```{r} 
#--- Mod3

# Chi-square and p-value 
 mod3_2step <- lm(ghg_log ~ gt_log_avg_pred2 + dt_log_avg + I(gt_log_avg_pred2 * dt_log_avg) + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)
 dat$residuals <- mod3_2step$residuals
 chisq <- summary(lm(residuals ~ qoi_log + extreme_left, dat))$r.squared * nrow(dat)
 chisq
 pchisq(chisq, 1, lower.tail=FALSE)
``` 


### Model implications 
See Figure 5. Marginal effects of environmental and digital technologies on GHG emissions

```{r} 
# Model estimates
 mod3_2step <- lm(ghg_log ~ gt_log_avg_pred2 + dt_log_avg + gt_log_avg_pred2:dt_log_avg + pop_log + va_ind_log + intermediate + rural + psm_ghg_log + year, dat = dat)

# Marginal effects of green technologies
 cplot(mod3_2step, x = "dt_log_avg", dx = "gt_log_avg_pred2", what = "effect",
      xlab = "Digital Tech (log)",
      ylab = "Marginal effect of Green Tech (log)"
      )
``` 


```{r} 
# Marginal effects of digital technologies
cplot(mod3_2step, x = "gt_log_avg_pred2", dx = "dt_log_avg", what = "effect",
     xlab = "Green Tech (log)",
     ylab = "Marginal effect of Digital Tech (log)",
     xlim = c(0, 7))
``` 


See Figure 6. Visualisation of the main effects of the “twin” transition
(a) Contour plot of predicted GHG emissions

```{r} 
 image(mod3_2step, "dt_log_avg", "gt_log_avg_pred2",
       xlab = "Digital Tech (log)", 
       ylab = "Green Tech (log)")
``` 

```{r}
# 3D plot (not shown in the paper)
 persp(mod3_2step, "dt_log_avg", "gt_log_avg_pred2", theta = 150, phi=0,
       xlab="Digital Tech (log)", ylab="Green Tech Instrum  (log)", zlab="GHG Emissions (log)", r=5)
```    

See Figure 6. Visualisation of the main effects of the “twin” transition
(b) Scatterplot of environmental and digital technologies

```{r}
 plot(dat$dt_log_avg, dat$gt_log_avg,
      xlab = "Digital Tech (log)", ylab = "Green Tech (log)")
 abline(0, 1, col = "red", lty = 2)
```    